\documentclass[11pt]{article}
\usepackage{MCATstyle}

\begin{document}
\section{BioChem}
\begin{itemize}
  \item The melting point of a molecule is defined to be the temperature at
    which 50\% of the molecules are denatured
  \item Guanine has a higher MW than adenosine
  \item A non-reducing SDS-PAGE will not break disulfide bonds.
  \item A reducing sugar is any sugar that is capable of acting as a reducing
    agent more specifically if a sugar has a hemiketal or hemiacetal group then
    it is a reducing sugar. All monosaccharides are reducing sugars.
    Disaccharides that have only one of their anomeric carbons participating in
    the glycosidic bonds are reducing e.g. lactose and maltose.
  \item Benedict's reagent tests for reducing sugars.
  \item Protein yield is best determined using active units i.e. protein weight
    times the specific activity. So don't just look at total protein purified
    look at that times the specific activity.
  \item The unfolding of proteins is a cooperative process and thus would appear
    as a sigmoidal curve in a plot of fraction of unfolded protein vs.
    denaturant concentration
  \item When removing electrons from transition metals to make cations, take
    from the S orbital first. So electron configuration of Co(II) is
    $ [Ar]3d^{7} $ as opposed to $ [Ar]4s^{2}3d^{5} $
  \item Increased conjugation means increased wavelength of light that is
    absorbed.
  \item Fructose is a ketose and the other two are aldose
  \item The start codon is AUG and the stop codons are UAA UGA UAG
  \item Totipotency is the ability of a cell to develop into any mature cell
    type
  \item Induction signaling occurs in development and is when a cell can induce
    changes in structure or function of neighboring cells.
\end{itemize}

\section{Psych}
\begin{itemize}
  \item Impression management: three selves
    \begin{itemize}
      \item authentic self describes who the person actually is
      \item Ideal self is who the person actually wants to be
      \item Tactical self is who we market our selves to be to adhere to the
        group's values
    \end{itemize}
  \item Drive reduction theory
    \begin{itemize}
      \item Primary drive are those essential for survival (innate)
      \item Secondary drive are those that are non-essential (learned by
        conditioning) such as the desire
        to succeed
    \end{itemize}
  \item Lateral geniculate nucleus responsible for visually processing shape and
    movement 
  \item Implicit memory of something can be implied through expression of
    emotion to that thing
  \item Four stages of demographic transition
    \begin{enumerate}
      \item high birth and death rates
      \item improvements in healthcare, sanitation, nutrition, and wages
      \item birth rates drop and economy shifts from agricultural to industrial
      \item low birth and death rates
    \end{enumerate}
  \item In times of stress hypothalamus releases CRF inducing ACTH release from
    anterior pituitary causing adrenal glands to release coritsol and cortisone.
  \item Histrionic personality disorder is characterized by constant
    attention-seeking behavior
  \item Stated characteristics of globalization:
    \begin{itemize}
      \item Rising levels of global unemployment
      \item Elevated pollution levels
      \item Decreased barriers to cultural exchange
    \end{itemize}
  \item Plutocracy is rule by upper classes, low social mobility, and uneven
    distribution of wealth
  \item Functions
    \begin{itemize}
      \item Manifest are conscious
      \item Latent are subconscious
    \end{itemize}
  \item \textbf{Anomie} refers to the breakdown of social bonds between an
    individual and society
  \item \textbf{Reliance on central traits} the tendency to organize the
    perception of others based on traits and personal characteristics of the
    target that are most important to the perceiver 
\end{itemize}


\end{document}
